---
title: "Credit Card Customer Analysis"
output: 
  html_document:
    keep_md: true
---

## Loading Packages

```{r package, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load required packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(knitr)
library(kableExtra)
```

## Data PreProcessing

### Basic Setup

original data retrieved from Kaggle <br>
try to understand data

read dataset.. remove weird columns

```{r setup, include=TRUE}
#read data
df <- read.csv("BankChurners.csv", stringsAsFactors = F)

#remove columns that contain "Naive"
df <- df %>%
  select(-contains('Naive'))

#check column names again
names(df)

summary_df <- describe(df)
summary_df
#unlist(summary_df[[1]])[2:12]

# check the number of unique client number
length(unique(df[,"CLIENTNUM"])) # total rows are 10127, so there is no duplicates

# Check for NA values in column
nrow(df[df$Avg_Utilization_Ratio == "NA",])
# need more of these!!!!!!!!!!
```
```{r outlier}
cols <- colnames(df)
cols <- cols[-1]
par(mfrow=c(4,5))
for (i in cols){
    plot <- ggplot(df, aes_string(x = df$CLIENTNUM, y = i)) +
    geom_point()
    print(plot)
}

# remove outlier
df1 <- df[!(df$Customer_Age %in% boxplot(df$Customer_Age)$out),]
df1 <- df[!(df$Total_Amt_Chng_Q4_Q1 %in% boxplot(df$Total_Amt_Chng_Q4_Q1)$out),]
df1 <- df[!(df$Total_Ct_Chng_Q4_Q1 %in% boxplot(df$Total_Ct_Chng_Q4_Q1)$out),]

#attach(df)
#par(mfrow=c(2,7))
#scatter1 <- plot(CLIENTNUM, Customer_Age)
#scatter2 <- plot(CLIENTNUM, Dependent_count)
#scatter3 <- plot(CLIENTNUM, Months_on_book)
#scatter4 <- plot(CLIENTNUM, Total_Relationship_Count)
#scatter5 <- plot(CLIENTNUM, Months_Inactive_12_mon)
#scatter6 <- plot(CLIENTNUM, Contacts_Count_12_mon)
#scatter7 <- plot(CLIENTNUM, Credit_Limit)
#scatter8 <- plot(CLIENTNUM, Total_Revolving_Bal)
#scatter9 <- plot(CLIENTNUM, Avg_Open_To_Buy)
#scatter10 <- plot(CLIENTNUM, Total_Amt_Chng_Q4_Q1)
#scatter11 <- plot(CLIENTNUM, Total_Trans_Amt)
#scatter12 <- plot(CLIENTNUM, Total_Trans_Ct)
#scatter13 <- plot(CLIENTNUM, Total_Ct_Chng_Q4_Q1)
#scatter14 <- plot(CLIENTNUM, Avg_Utilization_Ratio)
```

```{r remove}
cols <- colnames(df1)
cols <- cols[-1]
par(mfrow=c(4,5))
for (i in cols){
    plot <- ggplot(df1, aes_string(x = df1$CLIENTNUM, y = i)) +
    geom_point()
    print(plot)
}

#par(mfrow=c(2,7))
#attach(df1)
#scatter1 <- plot(CLIENTNUM, Customer_Age)
#scatter2 <- plot(CLIENTNUM, Dependent_count)
#scatter3 <- plot(CLIENTNUM, Months_on_book)
#scatter4 <- plot(CLIENTNUM, Total_Relationship_Count)
#scatter5 <- plot(CLIENTNUM, Months_Inactive_12_mon)
#scatter6 <- plot(CLIENTNUM, Contacts_Count_12_mon)
#scatter7 <- plot(CLIENTNUM, Credit_Limit)
#scatter8 <- plot(CLIENTNUM, Total_Revolving_Bal)
#scatter9 <- plot(CLIENTNUM, Avg_Open_To_Buy)
#scatter10 <- plot(CLIENTNUM, Total_Amt_Chng_Q4_Q1)
#scatter11 <- plot(CLIENTNUM, Total_Trans_Amt)
#scatter12 <- plot(CLIENTNUM, Total_Trans_Ct)
#scatter13 <- plot(CLIENTNUM, Total_Ct_Chng_Q4_Q1)
#scatter14 <- plot(CLIENTNUM, Avg_Utilization_Ratio)

```

```{r dddd, include=FALSE}
#boxplot showing outliers
par(mfrow=c(2,7))
boxplot(df$Customer_Age, ylab = "Customer_Age")
boxplot(df$Dependent_count, ylab = "Dependent_count")
boxplot(df$Months_on_book, ylab = "Months_on_book")
boxplot(df$Total_Relationship_Count, ylab = "Total_Relationship_Count")
boxplot(df$Months_Inactive_12_mon, ylab = "Months_Inactive_12_mon")
boxplot(df$Contacts_Count_12_mon, ylab = "Contacts_Count_12_mon")
boxplot(df$Credit_Limit, ylab = "Credit_Limit")
boxplot(df$Total_Revolving_Bal, ylab = "Total_Revolving_Bal")
boxplot(df$Avg_Open_To_Buy, ylab = "Avg_Open_To_Buy")
boxplot(df$Total_Amt_Chng_Q4_Q1, ylab = "Total_Amt_Chng_Q4_Q1")
boxplot(df$Total_Trans_Amt, ylab = "Total_Trans_Amt")
boxplot(df$Total_Trans_Ct, ylab = "Total_Trans_Ct")
boxplot(df$Total_Ct_Chng_Q4_Q1, ylab = "Total_Ct_Chng_Q4_Q1")
boxplot(df$Avg_Utilization_Ratio, ylab = "Avg_Utilization_Ratio")

# remove outliers
#df1 <- df[!(df$Customer_Age %in% boxplot(df$Customer_Age)$out),]

outlier <- boxplot(df$Customer_Age)$out
df1 <- df[!(df$Customer_Age %in% outlier),]
outlier <- boxplot(df$Months_on_book)$out
df1 <- df[!(df$Months_on_book %in% outlier),]
outlier <- boxplot(df$Months_Inactive_12_mon)$out
df1 <- df[!(df$Months_Inactive_12_mon %in% outlier),]
outlier <- boxplot(df$Contacts_Count_12_mon)$out
df1 <- df[!(df$Contacts_Count_12_mon %in% outlier),]
outlier <- boxplot(df$Credit_Limit)$out
df1 <- df[!(df$Credit_Limit %in% outlier),]
outlier <- boxplot(df$Avg_Open_To_Buy)$out
df1 <- df[!(df$Avg_Open_To_Buy %in% outlier),]
outlier <- boxplot(df$Total_Amt_Chng_Q4_Q1)$out
df1 <- df[!(df$Total_Amt_Chng_Q4_Q1 %in% outlier),]
outlier <- boxplot(df$Total_Trans_Amt)$out
df1 <- df[!(df$Total_Trans_Amt %in% outlier),]
outlier <- boxplot(df$Total_Trans_Ct)$out
df1 <- df[!(df$Total_Trans_Ct %in% outlier),]
outlier <- boxplot(df$Total_Ct_Chng_Q4_Q1)$out
df1 <- df[!(df$Total_Ct_Chng_Q4_Q1 %in% outlier),]

dev.new()
nrow(df1)
par(mfrow=c(2,7))
boxplot(df1$Customer_Age, ylab = "Customer_Age")
boxplot(df1$Dependent_count, ylab = "Dependent_count")
boxplot(df1$Months_on_book, ylab = "Months_on_book")
boxplot(df1$Total_Relationship_Count, ylab = "Total_Relationship_Count")
boxplot(df1$Months_Inactive_12_mon, ylab = "Months_Inactive_12_mon")
boxplot(df1$Contacts_Count_12_mon, ylab = "Contacts_Count_12_mon")
boxplot(df1$Credit_Limit, ylab = "Credit_Limit")
boxplot(df1$Total_Revolving_Bal, ylab = "Total_Revolving_Bal")
boxplot(df1$Avg_Open_To_Buy, ylab = "Avg_Open_To_Buy")
boxplot(df1$Total_Amt_Chng_Q4_Q1, ylab = "Total_Amt_Chng_Q4_Q1")
boxplot(df1$Total_Trans_Amt, ylab = "Total_Trans_Amt")
boxplot(df1$Total_Trans_Ct, ylab = "Total_Trans_Ct")
boxplot(df$Total_Ct_Chng_Q4_Q1, ylab = "Total_Ct_Chng_Q4_Q1")
boxplot(df1$Avg_Utilization_Ratio, ylab = "Avg_Utilization_Ratio")

#https://stackoverflow.com/questions/49044753/scale-kable-table-to-fit-page-width
#kable(head(df,5), format = "latex") %>%
#  kable_styling(latex_options = "scale_down")
#dt <- df[1:10,]
#dt <- dt %>%
#  kbl() %>%
#  kable_styling(latex_options = "scale_down") 
#dt
```

### Binning
Next step would be 'Binning' continuous variables into specific groups.
need some transformation before analysis

#### Age
Categorize age into 6 groups
```{r age, echo=TRUE}
min(df1$Customer_Age)
max(df1$Customer_Age)
age_labels <- c("20-29","30-39","40-49","50-59","60-69","70+")
df2 <- df1 %>%
  mutate(Age = cut(Customer_Age,
                   breaks = c(20,30,40,50,60,70,80), 
                   right = F, 
                   labels = age_labels)) 

#check the age binning by looking at randomly chosen data subset
age_check <- df2 %>%
  select(c(Customer_Age,Age)) %>%
  sample_n(10)

age_check

#remove continuous age value column
df2 <- df2 %>%
  relocate(Age, .after = Customer_Age) %>%
  select(-Customer_Age)

df2[1:5,]
```

#### Number of Dependents
Categorize number of dependent into 4 groups
```{r dependent, echo=TRUE}
min(df2$Dependent_count)
max(df2$Dependent_count)
dependent_labels <- c("0","1-2","3-4","5+")
df2 <- df2 %>%
  mutate(Dependent_Count = cut(Dependent_count,
                               breaks = c(0,1,3,5,10), 
                               right = F, 
                               labels = dependent_labels)) 

#check the dependent count binning by looking at randomly chosen data subset
dependent_check <- df2 %>%
  select(c(Dependent_count,Dependent_Count)) %>%
  sample_n(10)

dependent_check

#remove continuous dependent count value column
df2 <- df2 %>%
  relocate(Dependent_Count, .after = Dependent_count) %>%
  select(-Dependent_count)
df2[1:5,]
```

## Create Tables for Analysis

examine groups of people/etc.. by comparing quantitative values
```{r na, echo=TRUE}
# Check for NA values in column
#nrow(df[df$Avg_Utilization_Ratio == "NA",])
```

#### Gender and Age
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r genderage, echo=TRUE}
#group by biography info
#group by gender and age for utilization ratio comparison
summary <- df2 %>%
  group_by(Gender, Age) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio))
kable(summary)
```

#### Marital status
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r married, echo=TRUE}
#group by family info
#group by marital status for multiple comparisons
summary2 <- df2 %>%
  group_by(Marital_Status) %>%
  summarise(mean_inactive_months_count = mean(Months_Inactive_12_mon),
            mean_util_ratio = mean(Avg_Utilization_Ratio),
            mean_trans_count = mean(Total_Trans_Ct))
kable(summary2)
#married who has the highest utilization ratio are least likely to be inactive for 12 months
```

#### Dependents
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r dependents, echo=TRUE}
#group by number of dependents
summary3 <- df2 %>%
  group_by(Dependent_Count) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio), 
            mean_trans_count = mean(Total_Trans_Ct),
            mean_revolving_balance = mean(Total_Revolving_Bal))
kable(summary3)
```

#### Card Class
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r card, echo=TRUE}
#group by card info
#calculated revolving balance percentage for better comparison
summary4 <- df2 %>%
  group_by(Card_Category) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio),
            mean_credit_limit = mean(Credit_Limit), 
            mean_revolving_balance = mean(Total_Revolving_Bal),
            revolving_balance_perc = (mean_revolving_balance/mean_credit_limit)*100) %>%
    arrange(match(Card_Category, c("Blue","Silver","Platinum","Gold")))
kable(summary4)
```

#### Education Level
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r education, echo=TRUE}
#group by education level for comparison of credit limit and revolving balance
summary5 <- df2 %>%
  group_by(Education_Level) %>%
  summarise(mean_credit_limit = mean(Credit_Limit), mean_revolving_balance = mean(Total_Revolving_Bal)) %>%
  arrange(match(Education_Level, c("Doctorate","Post-Graduate","Graduate","College",
                                   "High School","Uneducated","Unknown")))
kable(summary5)
```

<br><br>


