---
title: "Credit Card Customer Analysis"
output: 
  html_document:
    keep_md: true
---

### Loading Packages

```{r package, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load required packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(knitr)
library(kableExtra)
```

### Basic Data Description

For the purpose of this project, a dataset that consists of credit card customer information has been retrieved from Kaggle.

```{r setup, include=TRUE}
df <- read.csv("BankChurners.csv", stringsAsFactors = F) #read data
length(df) # number of observations in original dataset

#remove columns that contain "Naive"
df <- df %>%
  select(-contains('Naive'))

length(df) #number of observations in new dataset

names(df) #check column names

summary(df) # summary statistics

#check the number of unique client number
length(unique(df[,"CLIENTNUM"])) # total rows are 10127, so there is no duplicates in client number

#check for the number of NA, 0, missing values for each column
sapply(df, function(x) sum(is.na(x) | x == 0 | x == ""))

```

The original dataset had 10127 rows and 23 columns. As suggested from the author of dataset, the last two columns containing the word "Naive" has been removed from the dataset, resulting a dataset with 10127 rows and 21 columns. There was no duplicated values in 'CLIENTNUM' column. A considerable amount of NA values has been found in a number of columns, and they will be pre-processed as we go along.

### Outlier Removal

This section will focus on removing outliers from the dataset to produce statistically significant results. Scatter plot for each column will be created to identify existing outliers.

```{r outlier}
cols <- colnames(df) # column names
cols <- cols[-1] # exclude 'CLIENTNUM' column from outlier removal process
par(mfrow=c(4,5))
for (i in cols){
  plot <- ggplot(df, aes_string(x = df$CLIENTNUM, y = i)) + geom_point()
  print(plot)
}
```

Looking at the scatter plot of each column, I could observe some outlier data points on a few columns - 'Customer_Age', 'Tota_Amt_Chng_Q4_Q1', 'Total_Ct_Chng_Q4_Q1'.

```{r removal, results = 'hide'}
# identify and remove outliers using box plots
df1 <- df[!(df$Customer_Age %in% boxplot(df$Customer_Age)$out),]
df1 <- df[!(df$Total_Amt_Chng_Q4_Q1 %in% boxplot(df$Total_Amt_Chng_Q4_Q1)$out),]
df1 <- df[!(df$Total_Ct_Chng_Q4_Q1 %in% boxplot(df$Total_Ct_Chng_Q4_Q1)$out),]
```

```{r remove}
cols <- colnames(df1) # column names
cols <- cols[-1] # exclude 'CLIENTNUM' column from outlier removal process
par(mfrow=c(4,5))
for (i in cols){
    plot <- ggplot(df1, aes_string(x = df1$CLIENTNUM, y = i)) +
    geom_point()
    print(plot)
}

#par(mfrow=c(2,7))
#attach(df1)
#scatter1 <- plot(CLIENTNUM, Customer_Age)
#scatter2 <- plot(CLIENTNUM, Dependent_count)
#scatter3 <- plot(CLIENTNUM, Months_on_book)
#scatter4 <- plot(CLIENTNUM, Total_Relationship_Count)
#scatter5 <- plot(CLIENTNUM, Months_Inactive_12_mon)
#scatter6 <- plot(CLIENTNUM, Contacts_Count_12_mon)
#scatter7 <- plot(CLIENTNUM, Credit_Limit)
#scatter8 <- plot(CLIENTNUM, Total_Revolving_Bal)
#scatter9 <- plot(CLIENTNUM, Avg_Open_To_Buy)
#scatter10 <- plot(CLIENTNUM, Total_Amt_Chng_Q4_Q1)
#scatter11 <- plot(CLIENTNUM, Total_Trans_Amt)
#scatter12 <- plot(CLIENTNUM, Total_Trans_Ct)
#scatter13 <- plot(CLIENTNUM, Total_Ct_Chng_Q4_Q1)
#scatter14 <- plot(CLIENTNUM, Avg_Utilization_Ratio)
```

```{r dddd, include=FALSE}
#delete this part?
#boxplot showing outliers
par(mfrow=c(2,7))
boxplot(df$Customer_Age, ylab = "Customer_Age")
boxplot(df$Dependent_count, ylab = "Dependent_count")
boxplot(df$Months_on_book, ylab = "Months_on_book")
boxplot(df$Total_Relationship_Count, ylab = "Total_Relationship_Count")
boxplot(df$Months_Inactive_12_mon, ylab = "Months_Inactive_12_mon")
boxplot(df$Contacts_Count_12_mon, ylab = "Contacts_Count_12_mon")
boxplot(df$Credit_Limit, ylab = "Credit_Limit")
boxplot(df$Total_Revolving_Bal, ylab = "Total_Revolving_Bal")
boxplot(df$Avg_Open_To_Buy, ylab = "Avg_Open_To_Buy")
boxplot(df$Total_Amt_Chng_Q4_Q1, ylab = "Total_Amt_Chng_Q4_Q1")
boxplot(df$Total_Trans_Amt, ylab = "Total_Trans_Amt")
boxplot(df$Total_Trans_Ct, ylab = "Total_Trans_Ct")
boxplot(df$Total_Ct_Chng_Q4_Q1, ylab = "Total_Ct_Chng_Q4_Q1")
boxplot(df$Avg_Utilization_Ratio, ylab = "Avg_Utilization_Ratio")

# remove outliers
#df1 <- df[!(df$Customer_Age %in% boxplot(df$Customer_Age)$out),]

outlier <- boxplot(df$Customer_Age)$out
df1 <- df[!(df$Customer_Age %in% outlier),]
outlier <- boxplot(df$Months_on_book)$out
df1 <- df[!(df$Months_on_book %in% outlier),]
outlier <- boxplot(df$Months_Inactive_12_mon)$out
df1 <- df[!(df$Months_Inactive_12_mon %in% outlier),]
outlier <- boxplot(df$Contacts_Count_12_mon)$out
df1 <- df[!(df$Contacts_Count_12_mon %in% outlier),]
outlier <- boxplot(df$Credit_Limit)$out
df1 <- df[!(df$Credit_Limit %in% outlier),]
outlier <- boxplot(df$Avg_Open_To_Buy)$out
df1 <- df[!(df$Avg_Open_To_Buy %in% outlier),]
outlier <- boxplot(df$Total_Amt_Chng_Q4_Q1)$out
df1 <- df[!(df$Total_Amt_Chng_Q4_Q1 %in% outlier),]
outlier <- boxplot(df$Total_Trans_Amt)$out
df1 <- df[!(df$Total_Trans_Amt %in% outlier),]
outlier <- boxplot(df$Total_Trans_Ct)$out
df1 <- df[!(df$Total_Trans_Ct %in% outlier),]
outlier <- boxplot(df$Total_Ct_Chng_Q4_Q1)$out
df1 <- df[!(df$Total_Ct_Chng_Q4_Q1 %in% outlier),]

dev.new()
nrow(df1)
par(mfrow=c(2,7))
boxplot(df1$Customer_Age, ylab = "Customer_Age")
boxplot(df1$Dependent_count, ylab = "Dependent_count")
boxplot(df1$Months_on_book, ylab = "Months_on_book")
boxplot(df1$Total_Relationship_Count, ylab = "Total_Relationship_Count")
boxplot(df1$Months_Inactive_12_mon, ylab = "Months_Inactive_12_mon")
boxplot(df1$Contacts_Count_12_mon, ylab = "Contacts_Count_12_mon")
boxplot(df1$Credit_Limit, ylab = "Credit_Limit")
boxplot(df1$Total_Revolving_Bal, ylab = "Total_Revolving_Bal")
boxplot(df1$Avg_Open_To_Buy, ylab = "Avg_Open_To_Buy")
boxplot(df1$Total_Amt_Chng_Q4_Q1, ylab = "Total_Amt_Chng_Q4_Q1")
boxplot(df1$Total_Trans_Amt, ylab = "Total_Trans_Amt")
boxplot(df1$Total_Trans_Ct, ylab = "Total_Trans_Ct")
boxplot(df$Total_Ct_Chng_Q4_Q1, ylab = "Total_Ct_Chng_Q4_Q1")
boxplot(df1$Avg_Utilization_Ratio, ylab = "Avg_Utilization_Ratio")

#https://stackoverflow.com/questions/49044753/scale-kable-table-to-fit-page-width
#kable(head(df,5), format = "latex") %>%
#  kable_styling(latex_options = "scale_down")
#dt <- df[1:10,]
#dt <- dt %>%
#  kbl() %>%
#  kable_styling(latex_options = "scale_down") 
#dt
```

### Binning

Some of the columns in the dataset consists of continuous variables, and that will need some transformation before further analysis. The following section will show the process of binning continuous variables into specific groups.


#### Age

The original 'Customer_Age' column is made of continuous age variables ranging from 20 to 80. In this section, the variable will be categorized into a total of 6 groups.

```{r age, echo=TRUE}
min(df1$Customer_Age) # minimum value of 'Customer_Age' column
max(df1$Customer_Age) # maximum value
age_labels <- c("20-29","30-39","40-49","50-59","60-69","70+") # define groups
df2 <- df1 %>% # break 'Customer_Age' column data into defined groups and store it as a new column
  mutate(Age = cut(Customer_Age,
                   breaks = c(20,30,40,50,60,70,80), 
                   right = F, 
                   labels = age_labels)) 

#check the age binning by looking at 10 randomly chosen data subset
age_check <- df2 %>%
  select(c(Customer_Age,Age)) %>%
  sample_n(10)

age_check # correctly done!

#remove continuous age value column
df2 <- df2 %>%
  relocate(Age, .after = Customer_Age) %>%
  select(-Customer_Age)

df2[1:5,] # check
```

#### Number of Dependents

This section will categorize the 'Dependent_count' column which consists of number of dependents data into 4 different groups. 
```{r dependent, echo=TRUE}
min(df2$Dependent_count) # min value of 'Dependent_count' column
max(df2$Dependent_count) # max value
dependent_labels <- c("0","1-2","3-4","5+") # define binning groups
df2 <- df2 %>% # break the column into defined groups of data and store it as a new column
  mutate(Dependent_Count = cut(Dependent_count,
                               breaks = c(0,1,3,5,10), 
                               right = F, 
                               labels = dependent_labels)) 

#check the dependent count binning by looking at randomly chosen data subset
dependent_check <- df2 %>%
  select(c(Dependent_count,Dependent_Count)) %>%
  sample_n(10)

dependent_check

df2 <- df2 %>% #remove continuous dependent count value column
  relocate(Dependent_Count, .after = Dependent_count) %>%
  select(-Dependent_count)
df2[1:5,] # check
```

### Create Tables for Analysis

The following section will examine and compare quantitative values, and look for any meaningful insights from the dataset.

```{r na, echo=TRUE}
# Check for NA values in column
#check for the number of NA, 0, missing values for each column
sapply(df, function(x) sum(is.na(x) | x == 0 | x == ""))

sapply(df2, function(x) sum(is.na(x) | x == 0 | x == ""))
```

#### Gender and Age

In this section, the dataset has been grouped by gender and age. I then calculated the mean utilization ratio for each group to compare card usage across diferent gender and age groups.

```{r genderage, echo=TRUE}
#group by biography info
#group by gender and age for utilization ratio comparison
summary <- df2 %>%
  group_by(Gender, Age) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio))
kable(summary)


```

#### Marital status
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r married, echo=TRUE}
#group by family info
#group by marital status for multiple comparisons
summary2 <- df2 %>%
  group_by(Marital_Status) %>%
  summarise(mean_inactive_months_count = mean(Months_Inactive_12_mon),
            mean_util_ratio = mean(Avg_Utilization_Ratio),
            mean_trans_count = mean(Total_Trans_Ct))
kable(summary2)
#married who has the highest utilization ratio are least likely to be inactive for 12 months
```

#### Dependents
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r dependents, echo=TRUE}
#group by number of dependents
summary3 <- df2 %>%
  group_by(Dependent_Count) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio), 
            mean_trans_count = mean(Total_Trans_Ct),
            mean_revolving_balance = mean(Total_Revolving_Bal))
kable(summary3)
```

#### Card Class
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r card, echo=TRUE}
#group by card info
#calculated revolving balance percentage for better comparison
summary4 <- df2 %>%
  group_by(Card_Category) %>%
  summarise(mean_util_ratio = mean(Avg_Utilization_Ratio),
            mean_credit_limit = mean(Credit_Limit), 
            mean_revolving_balance = mean(Total_Revolving_Bal),
            revolving_balance_perc = (mean_revolving_balance/mean_credit_limit)*100) %>%
    arrange(match(Card_Category, c("Blue","Silver","Platinum","Gold")))
kable(summary4)
```

#### Education Level
dlkfnalsdnflsnflansdlfnaflansflanlafnlanlnflnlna
```{r education, echo=TRUE}
#group by education level for comparison of credit limit and revolving balance
summary5 <- df2 %>%
  group_by(Education_Level) %>%
  summarise(mean_credit_limit = mean(Credit_Limit), mean_revolving_balance = mean(Total_Revolving_Bal)) %>%
  arrange(match(Education_Level, c("Doctorate","Post-Graduate","Graduate","College",
                                   "High School","Uneducated","Unknown")))
kable(summary5)
```




