---
title: "HR Analysis"
output: 
  html_document:
    keep_md: true
---
### Loading Packages

```{r package, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
#load required packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(knitr)
library(kableExtra)
library(scales)
```

### Data PreProcessing

#### Basic Data Information

```{r info, include=TRUE}
#read data
test <- read.csv("aug_test.csv", stringsAsFactors = F)
train <- read.csv("aug_train.csv", stringsAsFactors = F)

#column names
colnames(test)
colnames(train)

#summary statistics
summary_test <- describe(test)
summary_test
summary_train <- describe(train)
summary_train

#check for duplicate enrollee id
#data has duplicates if number of unique value = number of observations
length(unique(test[,"enrollee_id"])) # no duplicates
length(unique(train[,"enrollee_id"])) # no duplicates

#number of NA, 0, missing values for each column
sapply(test, function(x) sum(is.na(x) | x == 0 | x == ""))
sapply(train, function(x) sum(is.na(x) | x == 0 | x == ""))

#test1 <- test1[!(test1$Customer_Age %in% boxplot(df$Customer_Age)$out),]
```

### Basic Visualization 

#### Gender

```{r gender, include=TRUE}
sex <- train %>% # create a dataframe called sex
  filter(!(gender == "")) %>% # remove blank values from 'gender' column
  group_by(gender) %>% # group by gender
  summarise(percent = n()/nrow(.) * 100) # calculate gender ratios

# create pie chart for gender
pie <- ggplot(sex, aes(x = "", y = percent, fill = reorder(gender,percent))) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(x = 1.58, label = paste0(round(percent,1),"%")),
            position = position_stack(vjust = .5), size = 2.3) +
  coord_polar("y", start = 0) +
  labs(fill = "Gender") +
  guides(fill = guide_legend(reverse = TRUE)) # reverse order of legend
pie
```
여자보다 남자 비율이 월등히 높음.
실제를 반영한 것일 수도 있지만 그냥 데이터가 imbalanced한 것일 수도 있음을
생각해야함

#### Major

```{r major, include=TRUE}
# pie chart for major discipline
major <- train %>%
  filter(!(major_discipline == "")) %>% # remove blank values from 'major_discipline'
  group_by(major_discipline) %>%
  summarise(percent = n()/nrow(.) * 100)

pie <- ggplot(major, aes(x = "", y = percent, fill = reorder(major_discipline, percent))) +
  # pies in order of largest percentage to smallest percentage
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(x = 1.58, label = paste0(round(percent,1),"%")),
            position = position_stack(vjust = .5), size = 1.85) + # percentage labels
  coord_polar("y", start = 0) +
  labs(fill = "Major Discipline") +
  guides(fill = guide_legend(reverse = TRUE))
pie
```

#### Education Level 
```{r edu, include=T}
# pie chart for education level
education <- train %>%
  filter(!(education_level == "")) %>% # remove blank values from 'major_discipline'
  group_by(education_level) %>%
  summarise(percent = n()/nrow(.) * 100)

pie <- ggplot(education, aes(x = "", y = percent, fill = reorder(education_level,percent))) +
  # pies in order of largest percentage to smallest percentage
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(x = 1.6, label = paste0(round(percent,1),"%")),
            position = position_stack(vjust = .5), size = 2.3) +
  coord_polar("y", start = 0) +
  labs(fill = "Education Level") +
  guides(fill = guide_legend(reverse = TRUE)) # reverse order of legend
pie

#pie <- ggplot(major, aes(x = "", y = percent, fill = major_discipline)) +
#  geom_bar(width = 1, stat = "identity", color = "white") +
#  scale_y_continuous(breaks = round(cumsum(rev(major$percent)), 1)) +
#  geom_text(aes(label = paste(round(percent,1),"%")),position = position_stack(vjust = .5)) +
#  coord_polar("y", start = 0)
#pie
```

#### Years of Experience
```{r exp, include=T}
#my_order <- c('<1','1','2','3','4','5','6','7','8','9','10','11','12','13','14','15',
#           '16','17','18','19','20','>20')

#exp <- train %>%
#  filter(!(experience == "")) %>%
#  #mutate(experience = ifelse(experience == "<1", "0",experience)) %>%
#  #mutate(experience = as.numeric(experience)) %>%
#  group_by(experience) %>%
#  summarise(count = n()) %>%
#  arrange(match(experience, my_order))

exp <- train %>%
  filter(!(experience == "")) %>% # remove missing values from 'experience' column
  filter(!experience == ">20") %>% # remove >20 years of experience
  #filter(!experience == "<1") %>% # remove <1 years of experience
  mutate(experience = ifelse(experience == "<1", "0",experience)) %>% # treat <1 experience as 0 experience
  mutate(experience = as.numeric(experience)) %>% # convert to numeric variable
  group_by(experience) %>%
  summarise(count = n()) # count for each 'years of experience'

# create a bar graph for years of experience distribution
bar <- ggplot(exp, aes(x = experience, y = count)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) # count labels at the top of each bar
bar
```

peaks around 5, 10, and 15 years of experience
people tend to leave their company at that period of time?

#### Distribution of Training Hours by Company Type 
```{r comptype, include=T}
density <- train %>%
  filter(!company_type == "") %>% # remove missing values from 'company_type'
  filter(!training_hours == "") # remove missing values from 'training_hours"

# create a density plot of training hour distribution by company type
plot <- ggplot(density, aes(x = training_hours)) +
  geom_density(aes(color = company_type))
plot
```

### Visualization for Comparison
### (Looking for Company Change vs Those who are not)
#### Company Size
Is there any relationship between company size and willingness to leave company?
Are data sciensts more likely to be willing to change company if they are working for small company?
```{r size, include=T}
#my_order <- c('<10','10/49','50-99','100-500','500-999','1000-4999','5000-9999','10000+')

#size <- train %>%
#  filter(!(company_size == "")) %>%
#  group_by(company_size) %>%
#  summarise(count = n()) %>%
#  arrange(match(company_size, my_order))

#bar <- ggplot(size, aes(x = company_size, y = count)) +
#  geom_bar(fill = "#0073C2FF", stat = "identity") +
#  geom_text(aes(label = count), vjust = -0.3)
#bar

#-------
# looking for job change (target=1)
size <- train %>%
  filter(target == "1") %>% # filter data for people looking for job change
  filter(!company_size == "") %>%
  group_by(company_size) %>%
  summarise(percent = n()/nrow(.) * 100)

pie <- ggplot(size, aes(x = "", y = percent, fill = reorder(company_size,percent))) +
  # from largest percentage to smallest percentage
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(x = 1.6, label = paste0(round(percent,1),"%")),
            position = position_stack(vjust = .5), size = 2.3) +
  coord_polar("y", start = 0) +
  labs(fill = "Company Size", title = "Looking for Job Change") +
  guides(fill = guide_legend(reverse = TRUE)) # reverse order of legend
pie

# not looking for job change (target=0)
size <- train %>%
  filter(target == "0") %>% # filter data for people not looking for job change
  filter(!company_size == "") %>% # remove missing values
  group_by(company_size) %>%
  summarise(percent = n()/nrow(.) * 100) # calculate percentage

pie <- ggplot(size, aes(x = "", y = percent, fill = reorder(company_size,percent))) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(x = 1.6, label = paste0(round(percent,1),"%")),
            position = position_stack(vjust = .5), size = 2.3) +
  coord_polar("y", start = 0) +
  labs(fill = "Company Size") +
  guides(fill = guide_legend(reverse = TRUE)) # reverse order of legend
pie
```

#### Company Type
```{r type, include=T}
#total number
company <- train %>%
  filter(!company_type == "") %>%
  group_by(company_type) %>%
  summarise(count = n())

freq <- ggplot(company, aes(fct_rev(fct_reorder(company_type, count)),count)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) +
  labs(x = "Company Type")
freq

#looking for job change (target=1)
change <- train %>%
  filter(target == "1") %>%
  filter(!company_type == "") %>%
  group_by(company_type) %>%
  summarise(count = n(), percent = n()/nrow(.) * 100)

freq <- ggplot(change, aes(fct_rev(fct_reorder(company_type, count)),count)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = paste0(count," (", paste0(round(percent,1),"%"),")"), vjust = -0.3)) +
  labs(x = "Company Type", title = "Looking for Job Change")
freq

#not looking for job change (target=0)
nochange <- train %>%
  filter(target == "0") %>%
  filter(!company_type == "") %>%
  group_by(company_type) %>%
  summarise(count = n(), percent = n()/nrow(.) * 100)

freq <- ggplot(nochange, aes(fct_rev(fct_reorder(company_type, count)),count)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = paste0(count," (", paste0(round(percent,1),"%"),")"), vjust = -0.3)) +
  labs(x = "Company Type", title = "Not Looking for Job Change")
freq
```
#### Training Hours Distribution by groups of people
```{r train, include=T}
density <- train %>%
  filter(!company_type == "") %>% # remove missing values
  filter(!training_hours == "") %>% # remove missing values
  mutate(target = ifelse(target == "0", "Not Looking for Job Change","Looking for Job Change"))

plot <- ggplot(density, aes(x = training_hours)) +
  geom_density(aes(color = target))
plot
```

#### Years of Experience
```{r yearexp, include=T}
density <- train %>%
  filter(!experience == "") %>% # remove missing values
  filter(!experience == ">20") %>% # remove >20 experience
  filter(!experience == "<1") %>% # remove <1 experience
  mutate(experience = as.numeric(experience)) %>% # convert to numeric variables
  mutate(target = ifelse(target == "0", "Not Looking for Job Change","Looking for Job Change"))

plot <- ggplot(density, aes(x = experience)) +
  geom_density(aes(color = target)) +
  labs(x = "Experience (years)", y = "Density", fill = "Group")
plot
```

Those who are looking for job change tend to have less years of experience.

