---
title: "Netflix Analysis"
output: 
  html_document:
    keep_md: true
---
### Loading Packages

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
#load required packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)
library(rworldmap)
```

### Data Description

For the purpose of this project, a Netflix dataset sourced from Kaggle has been used.

```{r data, include = T}
df <- read.csv("netflix_titles.csv", stringsAsFactors = F) # read data

summary(df) # summary statistics
length(df$release_year) # number of observations

UniqueValue = function (x) {length(unique(x))} # shows number of unique values in each column
apply(df, 2, UniqueValue) # number of unique values in each column

sapply(df, function(x) sum(x == "")) # number of missing values
```

The original dataset consists of 7787 rows and 12 columns. All columns excluding 'show_id' and 'title' had unique values less than 7787, which is the total number of rows of this Netflix dataset. This indicates presence of duplicates in those columns. Also, many blank - "" - values has been found in 'director','cast','country','date_added', and 'rating' columns.

Cleaning all misleading values at once will lead to major data loss. In order to prevent data biases, pre-processing for this dataset will be done to a minimum. Cleaning will only be done on columns in use.

### Visualization

In this section, only relevant columns will be pre-processed, then be used for creating visualizations. Pre-processing methodologies will be discussed under each analysis topic. 

#### Netflix trends by year

This section will investigate the Netflix trend in their contents addition by year - regardless of contents release date. Note that a movie released in 2017 can be added on Netflix in 2020. How many movies and TV shows they add to their platform each year? Does Netflix treat movies and TV shows differently?

The 'date_added' column was originally in 'month day, year' format as character type. Therefore, it has been converted to numeric 'year' for yearly trend analysis. Missing values in 'date_added' also has been removed for the purpose of this analysis.

```{r year, include=T}
# To investigate yearly trends, I need to eliminate Month and Day data from 'date_added' column
head(df$date_added,5) # this column is originally in 'Month Day, Year' format
class(df$date_added) # column class: character

df$date_added <- sub(".*(\\d+{4}).*$", "\\1", df$date_added) # convert date to year
head(df$date_added,5) # this column now only has year data

# total number of movies and TV shows added to Netflix by year
year <- df %>%
  filter(!(date_added == "")) %>% # remove missing values
  group_by(date_added) %>%
  summarise(count = n()) # number of movies and TV shows by year added to Netflix
year

# draw a bar plot for number of contents added to Netflix by year
bar <- ggplot(year, aes(x = date_added, y = count, fill = date_added)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) +
  labs(x = "Year", fill = "Year", title = "Number of Contents added to Netflix by Year") + 
  theme(plot.title = element_text(hjust = 0.5)) # center title
bar

# Netflix trends - movie vs TV shows
group <- df %>%
  filter(!(date_added == "")) %>%
  group_by(type, date_added) %>% # group by content type and year added
  summarise(count = n())
group

# create bar plot for number of contents added by year by content type
bar <- ggplot(group, aes(x = date_added, y = count, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Netflix Trends by Year", x = "Year", "Type") + 
  theme(plot.title = element_text(hjust = 0.5)) # center title
bar
```
First bar plot shows the total number of movies and TV shows added on Netflix by year. After years of adding more contents each year, Netflix finally added less number of contents to its platform in 2020.

Second chart compares its trend in movies to that of TV shows. Although the number of movies added to Netflix outnumbers that of TV shows on Netflix, it started to show a decrease in 2020, whereas number of TV shows added has continued to increase. This may be an indication of Netflix focusing more on TV shows.

#### Movie vs TV show

This section will show total number of movies and TV shows on Netflix.

```{r type, include=T}
# create data frame for number of contents by type
type <- df %>%
  group_by(type) %>%
  summarise(count = n())
type

#all plots created afterwards will have title at center
#theme_update(plot.title = element_text(hjust = 0.5)) 

# create bar plot for Netflix contents by type
bar <- ggplot(type, aes(x = type, y = count, fill = type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) +
  labs(title = "Netflix contents by type", x = "Type", fill = "Type") + # add title
  theme(plot.title = element_text(hjust = 0.5)) # center title
bar

# create data frames each for movies and TV shows for further analysis
movie <- df[(df$type == "Movie"),]
tv <- df[(df$type == "TV Show"),]
```
As shown in the chart above, the number of movies on Netflix is more than double the number of TV shows.

For further analysis on each content type, the dataset has been split down into two separate datasets based on the 'type' column.

#### Netflix Movie Analysis

The remaining of this project will examine movies on Netflix. The dataset 'movie' that has been created above will be used.

#### Top 10 Countries by number of movies on Netflix

1) Bar plot
```{r country, include=T}
# basic information about movies on Netflix
length(unique(movie$country)) # country column has many duplicates
head(movie$country, 50) # some movies has listed multiple countries separated by comma

# split comma separated country names
country <- unlist(strsplit(movie$country, ", ")) # spliy country names at comma
unique(country) # some country names has comma at the end

# remove leftover commas from country names
country <- gsub(",","", country)
unique(country)

# store the list as data frame
movie_country <- data.frame(country = country)
unique(movie_country$country) # it now has clean country names

# count country names appearing in the original dataset
movie_country <- movie_country %>%
  group_by(country) %>%
  summarise(count = n())
head(movie_country,10)

# top 10 countries by number of movies added to Netflix
bar <- movie_country %>%
  arrange(desc(count)) %>% # most to least
  slice(1:10) %>% # top 10 countries
  ggplot(., aes(x=reorder(country,-count), y = count)) + # bar plot most to least
  geom_bar(stat='identity') + 
  theme_classic() +
  labs(x = "Country", title = "Top 10 Countries by number of movies on Netflix") +
  geom_text(aes(label = count), vjust = -0.3) # add count labels
bar
```

United States has the most number of movies on Netflix.

2) World Map
```{r map, include, results='hide'}
spdf <- joinCountryData2Map(movie_country, joinCode="NAME", nameJoinColumn="country")
mapParams <- mapCountryData(spdf,
               nameColumnToPlot="count",
               catMethod=c(0,1,3,5,10,50,100,300,500,1000,2500),
               mapTitle = "Number of Movies added to Netflix",
               addLegend = FALSE)
do.call(addMapLegend, c(mapParams, legendLabels="all", legendWidth=0.5))

#labelCountries()
```

Continent-wise, Americas - North and South America - exhibits high number of Netflix movies overall.

#### Number of movies by Year
year 지날때마다 각 나라의 영화 개수
line graph - x = year, y = count, fill = country

```{r countbyyear}
genre_country <- movie %>% 
  mutate(listed_in = strsplit(as.character(listed_in), ", ")) %>% # separate genre by commas
  unnest(listed_in) %>%
  mutate(country = strsplit(as.character(country), ", ")) %>% # separate country name by commas
  unnest(country) %>%
  select("title","country","date_added","listed_in")
genre_country

unique(genre_country$country) # some country names still have commas
length(unique(genre_country$country)) # number of unique country names
unique(genre_country$listed_in) # clean; no leftover commas to be deleted

genre_country <- genre_country %>%
  mutate(country = gsub(",","",country)) # remove commas by replacing them with space
genre_country

unique(genre_country$country) # commas removed from country names
length(unique(genre_country$country))

movie_year <- genre_country %>%
#  mutate(date_added = as.numeric(date_added)) %>%
  filter(country == "United States" |
           country == "India"| 
           country == "United Kingdom"| 
           country == "Canada"| 
           country == "France") %>%
  group_by(country, date_added) %>%
  summarise(count = n())
movie_year

#bar <- ggplot(genre_5_country, aes(fill=listed_in, y=count, x=country)) + 
#    geom_bar(position="fill", stat="identity")
#bar

chart <- ggplot(movie_year, aes(x = date_added, y = count, group = country, color = country)) +
  geom_line() +
  labs(x = "Year", y = "Count", fill = "Country", title = "Number of movies by Year in Top 5 Countries")
chart <- ggplotly(chart)
chart
```


#### Netflix movie genre trends by year
 장르 ('listed in' column) 별로 몇년도에 넷플에 애드 됐는지? (release year아니고 넷플 애드된 해)
 몇년도에 어떤 장르가 제일 많이 애드됐는지

```{r genrebyyear, include=T}
head(movie$listed_in,10) # each movie has multiple genres separated by commas

movie_genre <- movie %>% 
  mutate(listed_in = strsplit(as.character(listed_in), ", ")) %>% # split genre column by comma
  unnest(listed_in) %>%
  select("title","date_added","listed_in")
movie_genre

unique(movie_genre$listed_in) # all Netflix genre

movie_genre <- movie_genre %>%
  group_by(date_added, listed_in) %>%
  summarise(count = n()) # count number of genres by year
movie_genre

# stacked bar chart
bar <- ggplot(movie_genre, aes(fill=listed_in, y=count, x=date_added)) + 
  geom_bar(position="stack", stat="identity") +
  labs (x = "Year", fill = "Genre", title = "Genre by Year")
bar <- ggplotly(bar)
bar # interactive로 해서 한 장르의 시간별 변화도 관찰 가능

# percent stacked bar chart
bar <- ggplot(movie_genre, aes(fill=listed_in, y=count, x=date_added)) + 
  geom_bar(position="fill", stat="identity") +
  labs (x = "Year", y = "Percent", fill = "Genre", title = "Genre by Year")
bar
```

마지막 percent stacked barchart를 보면 해가 지날수록 장르가 더 다양해 지는 걸 볼 수 있었음
2014년을 기점으로 international과 LGBTQ movie의 비중이 대폭 늘어남
전체적으로 시간이 지날 수록 영화의 다양성에 집중하고자 하는 양상이 드러남

#### Netflix genre trends by country
```{r genrebycountry, include = TRUE}


#넷플 영화 갯수 탑 5 나라의 장르 비율
#genre_5_country <- genre_country %>%
#  filter(country == "United States" |
#           country == "India"| 
#           country == "United Kingdom"| 
#           country == "Canada"| 
#           country == "France") %>%
#  group_by(country, listed_in) %>%
#  summarise(count = n())

#  stacked bar chart
#bar <- ggplot(genre_5_country, aes(fill=listed_in, y=count, x=country)) + 
#    geom_bar(position="fill", stat="identity")
#bar

# 각 나라 별 도넛 차트 for genre
# https://homepage.divms.uiowa.edu/~luke/classes/STAT4580/catone.html
#pie <- ggplot(genre_5_country) +
#  geom_col(aes(x = 1, y = count, fill = listed_in), position = "fill") +
#  coord_polar(theta = "y") +
#  labs(title = "Movie Genres", fill = "Genre") +
#  facet_wrap(~ country) +
#  theme_bw() +
#  theme(axis.title = element_blank(), # remove axis and grid lines
#        axis.text = element_blank(),
#        axis.ticks = element_blank(),
#        panel.grid.major = element_blank(),
#        panel.grid.minor = element_blank(),
#        panel.border = element_blank())

#pie <- pie + xlim(0, 1.5)
#pie

# 영화 갯수 제일 많은 미국 한 나라의 장르 비율 도넛 차트
genre_usa <- genre_country %>%
  filter(country == "United States") %>%
  group_by(country, listed_in) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# donut chart 만들기 (많은순으로 정렬)
pie <- ggplot(genre_usa) +
  geom_col(aes(x = 1, y = count, fill = reorder(listed_in,count)), position = "fill") +
  coord_polar(theta = "y", start = 0) +
  labs(title = "Movie Genres in United States", fill = "Genre") +
  guides(fill = guide_legend(reverse = TRUE)) +
  facet_wrap(~ country) +
  theme_bw() +
  theme(axis.title = element_blank(), # remove axis and grid lines
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())

pie <- pie + xlim(0, 1.5)
pie

```

Very small portion of American movies are categorized as international movies. <- dㅣ건 미국 도넛차트만 둘거면 빼야할듯


